---
title: "Co-Kriging with GAM"
author: "YX"
date: "8/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r }

#list_libs <- c("rgdal","readr")
#lapply(list_libs, require, character.only = TRUE)

library(rgdal)
library(readr)
library(smerc)                                 #Implements statistical methods for analyzing the counts of areal data, with a 
library(maptools)
library(sf)
library(raster)
library(rasterVis)                        # The rasterVis package complements the raster package, providing a set of methods for enhanced              
library(dismo)
library(spatial)
library(sp)
library(maxstat)
library(rgdal)
library(RColorBrewer)
library(LaplacesDemon)
library(imager)
library(pracma)
library("rgeos")
library(maps)
library("rnaturalearth")
library("rnaturalearthdata")
library(ggspatial)
library(ggplot2)
library(wesanderson)
library(mltools)
library(gganimate)                           # create anamination
library(tools)
library(zoo)
library(lubridate)
library(ggnewscale)                          #use multiple colour and fill scales in ggplot2
library(gstat)
library(taRifx)
```


```{r }
## Read in the water level .csv file
path_waterlevel = ("input/data/water_levels")
water_level <- read.csv(paste(path_waterlevel,"/Shallow.Combined.Master_WellInfo_w_ShallowWLdata_1Yr_wBRAA_v2.csv",sep=''))
#water_level <- read.csv(paste(path_waterlevel,"/test_no_Jackson.csv",sep=''))


thre_cutoff <- as.integer(200000)
thre_width <- as.integer(15000)
#cal_year <- "Shallow.Combined.1929.noJackson" # WL2018, WL2019, WL2010, Shallow.Combined.1929, Shallow.Below.1929, 
                    # Shallow.Combined.1929.noJackson, Shallow.Below.1929.noJackson,   
#cal_year <- "WL2018"
#cal_year <- "WL2019"
#cal_year <- "WL2020"
#cal_year <- "Shallow.Combined.1929"
#cal_year <- "Shallow.Below.1929"

#cal_year <- "WL2018_minus_Shallow.Combined.1929"
#cal_year <- "WL2019_minus_Shallow.Combined.1929"
#cal_year <- "WL2020_minus_Shallow.Combined.1929"
#cal_year <- "WL2018_minus_Shallow.Below.1929"
#cal_year <- "WL2019_minus_Shallow.Below.1929"
cal_year <- "WL2020_minus_Shallow.Below.1929"



#list_cal_years <- list("WL2018","WL2019")


#for ( cal_year in list_cal_years){ 
# print(cal_year)
 




## read in the spatial polygons for each acquifer
path <- ("input/GIS/shapefiles/geology/District_Geology_Boundaries")

Yegua_Jackson <- readOGR(dsn = path, layer = "YJ_400_original")
Sparta        <- readOGR(dsn = path, layer = "SP_400_original")
Queen_City    <- readOGR(dsn = path, layer = "QC_400_original")
Carrizo       <- readOGR(dsn = path, layer = "CZ_400_original")
Calvert_Bluff <- readOGR(dsn = path, layer = "CB_400_original")
Simboro       <- readOGR(dsn = path, layer = "SB_400_original")
Hopper        <- readOGR(dsn = path, layer = "HP_400_original")
ShallowZone   <- readOGR(dsn = path, layer = "Total_400_original")

crs(ShallowZone) <- crs(Hopper)

## read in the spatial line df for river
path_river = ("input/GIS/shapefiles/hydrology")
river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_pts")

## read the DEM .tif file
DEM <- raster("input/GIS/rasters/DEM_POSGCD_NeighboringCounties.tif")

# import the pumping raster (difference in water level after pumping )
pumping500_2018 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2018.500ft.pst15.tif")
pumping500_2019 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2019.500ft.pst15.tif")
pumping500_2020 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")

pumping <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")

#plot(pumping500_2018)
#plot(pumping500_2020)

# import the GAM (simulated water levels) 
sim_WL2020 <- raster("input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2020.500ft.pst15.tif")
#sim_WL2019 <- raster("input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2019.500ft.pst15.tif")
plot(sim_WL2020)
#plot(sim_WL2019)


#setwd("c:/Users/hpham/OneDrive - INTERA Inc/projects/012_rscript_geostats_s/scripts/")

```

# create grids for kriging interpolations 
```{r}
# x <- seq(5800000, 6100000,400)
# y <- seq(19333000, 19633000,400)

x <- seq(5749040.9338320437818766, 6159040.9338320437818766,500)
y <- seq(19251409.7630450688302517, 19747409.7630450688302517,500)
data.grid <- meshgrid(x,y)

data.grid <- data.frame(unlist(matrix(data.grid$X)), unlist(matrix(data.grid$Y)))
colnames(data.grid) <- c("x","y")
data.grid$DEM = 0
coordinates(data.grid) <- ~ x + y
crs(data.grid) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
data.grid
```

# crop DEM topography raster to the same extent as the river and measured water levels 
```{r}
# create a empty raster to store topography DEM 
data.raster <- rasterFromXYZ(data.frame(data.grid@coords), res=c(500,500))
crs(data.raster) = CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
bb =raster::extent(5749040.9338320437818766, 6159040.9338320437818766, 19251409.7630450688302517, 19747409.7630450688302517) 
#c(5800000, 6100000,19333000,19633000)    #  6159040 - 5749040 = 410000 / 500 = 820   dimension of raster
data.raster[] = 0                         #  19747409 - 19251409 = 496000 / 500 = 992
data.raster

# DEM_raster <- raster(ncols = 750, nrows= 750)
# crs(DEM_raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
# extent(DEM_raster) <- extent(data.grid)
# DEM.raster <- resample(DEM, DEM_raster, method = "bilinear")
# DEM.raster
    
```

# Data wrangling, select specific cols for subsequent analysis and convert to spatial point data frame with correct CRS
```{r}
library(dplyr)
library(pracma)

# choose the year of WL you want to work on
# filter out NAs in both WL2019 and WL2020 to keep same number of measured water levels for both years
data <- water_level %>%
  dplyr::select(Longitude, Latitude, SWN, DEM, Depth, ScrTop1, ScrBot1, ScrTop2,
                ScrBot2, SH_Note, Aq2Use, cal_year)
# Shallow.Combined.1929, Shallow.Below.1929
#Shallow.Combined.1929.noJackson, Shallow.Below.1929.noJackson
#data$WL2015


#%>%   filter(cal_year != 0) 

str(data)

# we can also only exclude NAs for individual years
#data_2019 <- data %>% 
#  dplyr::select(Longitude, Latitude, WL2019Loc.DD.Adjt) %>%
#  filter(WL2019Loc.DD.Adjt != 0)
#
#data_2020 <- data %>% 
#  dplyr::select(Longitude, Latitude, cal_year) %>%
#  filter(WL2015 != 0)

#data_tmp <- data %>% 
#  dplyr::select(Longitude, Latitude, cal_year)

#data_tmp

data_2020 <- data %>% 
  dplyr::select(Longitude, Latitude, cal_year) %>%
  filter(data[,cal_year] != 0)

#data_2020[,cal_year]
#data_2020$WL2015

# create spatial point data frame with geographic coordinate systems
#data_sp2019 <- SpatialPointsDataFrame(SpatialPoints(data_2019[,1:2]), data.frame(data_2019[,3]))
#crs(data_sp2019) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

data_sp2020 <- SpatialPointsDataFrame(SpatialPoints(data_2020[,1:2]), data.frame(data_2020[,3]))
crs(data_sp2020) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

# create spatial point data frames with projected coordinate systems
new_crs <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
#data_sp2019 <- spTransform(data_sp2019,new_crs)
data_sp2020 <- spTransform(data_sp2020,new_crs)

# rename the colnames
#colnames(data_sp2019@coords)[1:2] = c("x","y")
#colnames(data_sp2019@data)        = c("WL2019")  
colnames(data_sp2020@coords)[1:2] = c("x","y")
colnames(data_sp2020@data)        = c("WL2020")  

# create combined spatial point data frame for both years (2019 & 2020) with projected coordinate systems
#data_spadf <- SpatialPointsDataFrame(SpatialPoints(data[,1:2]),data=data[,12:13])
#crs(data_spadf) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
#data_spadf <- spTransform(data_spadf,new_crs)
#data_spadf@bbox <- t(matrix(extent(data.grid),2))  #crop


#Normalized Data
normalized = function (x) {(x-min(x))/(max(x)-min(x))}                              # create a normalization function
#data_sp2019@data$WL2019n <- normalized(data_sp2019@data$WL2019)
data_sp2020@data$WL2020n <- normalized(data_sp2020@data$WL2020)
#data_spadf@data$WL2019n  <- normalized(data_spadf@data$WL2019)
#data_spadf@data$WL2020n  <- normalized(data_spadf@data$WL2020)
#str(data_sp2019@data)


```
# preprocessing the GAM (simulated water levels), and extract a subset spatialpointdataframe to variogram and co-krige with measured water levels
```{r}
# convert the selected GAM raster for either 2019 or 2020 to a spatial point data frame, specify the year of the data
sim_wl.data <- rasterToPoints(sim_WL2020, spatial = TRUE, progress = "window")
#str(sim_wl.data)

# rename the colname
colnames(sim_wl.data@data)[1] = "sim_wl"
# normalized the data
sim_wl.data@data$sim_wln <- normalized(sim_wl.data@data$sim_wl)
sim_wl.data
#str(sim_wl.data)
summary(sim_wl.data)

# convert the spatial point df to a regular dataframe
sim_wl.dataframe <- as.data.frame(sim_wl.data)
colnames(sim_wl.dataframe) <- c("sim_wl","sim_wln", "x", "y")

#str(sim_wl.dataframe)

# subset 1 in every 500 data for variogram modeling
sim_wl.dataframesubset <- subset(sim_wl.dataframe[seq(1,nrow(sim_wl.dataframe), 500),])
str(sim_wl.dataframesubset)
head(sim_wl.dataframesubset)

# create a subset DEM spatialpointdataframe for variogram modeling
sim_wl2020.datasubset <- SpatialPointsDataFrame(SpatialPoints(sim_wl.dataframesubset[,3:4]),data=data.frame(sim_wl.dataframesubset[,1:2]))
crs(sim_wl2020.datasubset) <- crs(data_sp2020)

#sim_wl2019.datasubset <- SpatialPointsDataFrame(SpatialPoints(sim_wl.dataframesubset[,3:4]),data=data.frame(sim_wl.dataframesubset[,1:2]))
#crs(sim_wl2019.datasubset) <- crs(data_spadf)

#summary(sim_wl2019.datasubset)
head(sim_wl2020.datasubset)
```

# Modelling the spatial structure of the target variable - make variogram
# 1. the variogram method computes the experimental variogram as a variogram object
# 2. the vgm method creates a variogram model object
# 3. the fit.variogram method adjusts a variogram model object to a variogram object

```{r}
# look at the spatial point data frame
#data_spadf
#str(data_spadf)


# plot the empirical variogram for measured water levels at specific years
v_WL2020n <- variogram(WL2020n ~ 1 , data=data_sp2020, cutoff=thre_cutoff, width=thre_width)
#str(v_WL2020n)
#summary(v_WL2020n)
#v_WL2020n$np
#v_WL2020n$dist

# Plot empirical variogram
#plot(v_WL2020n, main = "Empirical Variogram",pl = T)

# create a model for the empirical variogram
variogram_model_WL2020n <- vgm(0.05,"Gau",100000,0.003)
#str(variogram_model_WL2020n)
#summary(variogram_model_WL2020n)

# plot both the fitted variogram model with the empirical variogram
#plot(v_WL2020n, pl=T, model=variogram_model_WL2020n, main = "Variogram Model")
#(variogram_model_WL2020n_fit <- fit.variogram(v_WL2020n, variogram_model_WL2020n))

variogram_model_WL2020n_fit <- fit.variogram(v_WL2020n, variogram_model_WL2020n)
#str(variogram_model_WL2020n_fit)
#summary(variogram_model_WL2020n_fit)

# export fig to png
ofile = paste("output/fitted_variogram_",cal_year,"_cutoff_",toString(thre_cutoff),".png")

png(ofile,width=10,height=8,units="in",res=300)
plot(v_WL2020n, pl=T, model=variogram_model_WL2020n_fit, main = paste("Fitted OK Variogram Model ", cal_year,"_cutoff_",toString(thre_cutoff)))
dev.off()

# Save to a csv file
ofile = paste("output/fitted_variogram_",cal_year,"_cutoff_",toString(thre_cutoff),".csv")
write.csv(v_WL2020n,ofile)
```



# Co-Kriging with GAM (simulated water levels)
# Model the spatial structure of a covariate and its covariance with the target variable. This is called co-regionalisation.

```{r}
# plot the empirical variogram for simulated water levels, change the dataset to model the varigram for a different year
plot(v.sim_wln <- variogram(sim_wln ~ 1, sim_wl2020.datasubset, cutoff=300000, width=20000), main = "Empirical Variogram",pl = T)

# change width from 20,000 to 1,000 to change bin size
#plot(v.sim_wln <- variogram(sim_wln ~ 1, sim_wl2020.datasubset, cutoff=300000, width=1000), main = "Empirical Variogram",pl = T)

plot(v.sim_wln, pl = T)

# fit the variogram model
m.sim_wln <- vgm(.04, "Gau",150000,0.001)
# examine the fitted variogram model
(m.sim_wln.f <- fit.variogram(v.sim_wln, m.sim_wln))
plot(v.sim_wln, pl=T, model=m.sim_wln.f, main = "Variogram MOdeling on GAM")

#head(m.sim_wln)

```


# Building a data structure to model co-regionalisation
# We have to fit models to both the direct and cross-variograms simultaneously, 
# and these models must lead to a positive definite cokriging system.
# Fit a linear model of co-regionalisation: all models (direct and cross) have 
# the same shape and range, but may have different partial sills and nuggest.
```{r}
# create a model of co-regionalisation, be sure to check the years for both datasets
# mea wl data_sp2020
(g <- gstat(NULL, id = "WL", form = WL2020n ~ 1, data=data_sp2020))
summary(g)
str(g)
#print(data_spadf)
#print(sim_wl2020.datasubset)
print(data_sp2020)

# 
(g <- gstat(g,  id = "SIM_WL", form = sim_wln ~ 1, data=sim_wl2020.datasubset))

# Compute and display the two direct variogram and one cross-variogram
v.cross <- variogram(g)
str(v.cross)
plot(v.cross, pl=T, main = "Cross-Variogram on WL and GAM")
write.csv(v.cross,'v.cross.csv')
```

# Fitting a linear model of co-regionalisation
```{r}
(g <- gstat(g, id = "WL", model = m.sim_wln.f, fill.all=T))

# To ensure a positive definite co-krigingg system, use fit.lmc (fit linear model of co-regionalization)
# This takes the initial estimate, fits all the variograms, and then each of the partial sills is adjusted
(g <- fit.lmc(v.cross, g, fit.method=1, correct.diagonal=1.01))
print(g)
# examine the model fit for all variogram models
plot(variogram(g), model=g$model, main = "Cross Variogram Models")

```


# Perform co-kriging with simulated water levels
```{r}
# use the predict function from gstat to interpolate the co-kriging estimates
coK_sim <- predict(g, data.grid)
coK_sim

# constrain the predicted values strictly between 0 and 1
coK_sim@data$SIM_WL.pred <- ifelse(coK_sim@data$SIM_WL.pred >= 1.0, 1, coK_sim@data$SIM_WL.pred)
coK_sim@data$SIM_WL.pred <- ifelse(coK_sim@data$SIM_WL.pred <= 0.0, 0, coK_sim@data$SIM_WL.pred)

# find the min and max of the simulated water levels in the original scale
min <- min(sim_wl2020.datasubset@data$sim_wl)
max <- max(sim_wl2020.datasubset@data$sim_wl)

# back transform to original scales
InvNormalize <- function (x) {x * (max - min) + min}
coK_sim@data$SIM_WLn.pred = InvNormalize(coK_sim@data$SIM_WL.pred)

sp::spplot(coK_sim, "SIM_WLn.pred", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Co-Kriging Estimates of WL with simulated WL2019", xlab = "x", ylab = "y", widths = 10)
```


# extract values for each acquifer for cokriging with GAM
```{r}
# extract values from spatial polygons
Yegua_Jackson_df      <- as.data.frame(coK_sim[Yegua_Jackson,])
Sparta_df             <- as.data.frame(coK_sim[Sparta,])
Queen_City_df         <- as.data.frame(coK_sim[Queen_City,])
Carrizo_df           <- as.data.frame(coK_sim[Carrizo,])
Calvert_Bluff_df      <- as.data.frame(coK_sim[Calvert_Bluff,])
Simboro_df        <- as.data.frame(coK_sim[Simboro,])
Hopper_df             <- as.data.frame(coK_sim[Hopper,])
ShallowZone_df        <- as.data.frame(coK_sim[ShallowZone,])
min <- min(data_spadf@data$WL2019)
max <- max(data_spadf@data$WL2019)

# back transform to original scales
InvNormalize <- function (x) {x * (max - min) + min}

# find one average value for each acquifer
tol_avg   <- mean(coK_sim$SIM_WLn.pred)
YJ_avg    <- mean(Yegua_Jackson_df$SIM_WLn.pred)
SPA_avg   <- mean(Sparta_df$SIM_WLn.pred)
QC_avg    <- mean(Queen_City_df$SIM_WLn.pred)
CAR_avg   <- mean(Carrizo_df$SIM_WLn.pred)
CB_avg    <- mean(Calvert_Bluff_df$SIM_WLn.pred)
SIM_avg   <- mean(Simboro_df$SIM_WLn.pred)
HOP_avg   <- mean(Hopper_df$SIM_WLn.pred)
SHAZ_avg  <- mean(ShallowZone_df$SIM_WLn.pred)
CO_avg <- data.frame(name = c("YJ", "SPA", "QC", "CAR", "CB", "SIM", "HOP", "TOL", "SHAZ"), avg = c(YJ_avg, SPA_avg, QC_avg, CAR_avg, CB_avg, SIM_avg, HOP_avg, tol_avg, SHAZ_avg))
CO_avg

```

# save file and write on disk
```{r}
# create raster and store it
coK_sim@data$SIM_WL.predN <- InvNormalize(coK_sim$SIM_WLn.pred)
#k.cc
coK_sim
k.temp <- coK_sim[,-(1:5)]
k.temp
# ck.raster <- raster(nrows = 10000,ncols =10000)
# crs(ck.raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
# extent(ck.raster) <- extent(ko)
# ck.raster <- rasterize(k.c, ck.raster, field = "WL2020.predN", fun = mean)
# ck.raster
#writeRaster(ck.raster, "CKDEM_raster_2020", "GTiff", overwrite = TRUE)
# specify the image path on disk
image_path = "output/Output Shapefile"
writeOGR(obj = k.temp, layer = "CK_SIM_2020", dsn = image_path, driver = "ESRI Shapefile", overwrite_layer = TRUE)

```
