---
title: "Co-Kriging with Topography DEM"
author: "Yuchen Xiao"
date: "8/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rgdal)
library(smerc)                                 # Implements statistical methods for analyzing the counts of areal data, with a 
library(maptools)
library(sf)
library(raster)
library(rasterVis)                             # The rasterVis package complements the raster package, providing a set of methods for enhanced         
library(dismo)
library(spatial)
library(sp)
library(maxstat)
library(rgdal)
library(RColorBrewer)
library(LaplacesDemon)
library(imager)
library(pracma)
library(maps)
library(viridis)
library("rnaturalearth")
library("rnaturalearthdata")
library(ggspatial)
library(ggplot2)
library(wesanderson)
library(mltools)
library(RColorBrewer)
library(gganimate)                             # create anamination
library(tools)
library(maps)
library(zoo)
library(lubridate)
library(ggnewscale)                            # use multiple colour and fill scales in ggplot2
library(gstat)
library(taRifx)

## Read in the water level .csv file
path_waterlevel = ("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data")
water_level <- read.csv(paste(path_waterlevel,"/Master_WellInfo_w_ShallowWLdata_1Yr_wBRAA_v2.csv",sep=''))

## read in the spatial polygons for each acquifer
path = ("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/geology/District_Geology_Boundaries")
Yegua_Jackson <- readOGR(dsn = path, layer = "YJ_400_original")
Sparta        <- readOGR(dsn = path, layer = "SP_400_original")
Queen_City    <- readOGR(dsn = path, layer = "QC_400_original")
Carrizo       <- readOGR(dsn = path, layer = "CZ_400_original")
Calvert_Bluff <- readOGR(dsn = path, layer = "CB_400_original")
Simboro       <- readOGR(dsn = path, layer = "SB_400_original")
Hopper        <- readOGR(dsn = path, layer = "HP_400_original")
ShallowZone   <- readOGR(dsn = path, layer = "Total_400_original")
crs(ShallowZone) <- crs(Hopper)

## read in the spatial line df for river
path_river = ("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/hydrology")
#river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_GAM")
river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_pts_clip")

## read the DEM .tif file
# cliped DEM .tif file
DEM <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/rasters/500ft_res/DEMs/DEM_ShallowArea_Resample500.tif")
#DEM <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/rasters/DEM_POSGCD_NeighboringCounties.tif")


# import the pumping raster (difference in water level after pumping )
pumping500_2018 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2018.500ft.pst15.tif")
pumping500_2019 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2019.500ft.pst15.tif")
pumping500_2020 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")

# import the GAM (simulated water levels) 
sim_WL2020 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2020.500ft.pst15.tif")
sim_WL2019 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2019.500ft.pst15.tif")



setwd("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/R Code")

```

# create grids for kriging interpolations 
```{r}
# Ross
x <- seq(5749040.9338320437818766, 6159040.9338320437818766,500)
y <- seq(19251409.7630450688302517, 19747409.7630450688302517,500)
data.grid <- meshgrid(x,y)

data.grid <- data.frame(unlist(matrix(data.grid$X)), unlist(matrix(data.grid$Y)))
colnames(data.grid) <- c("x","y")
data.grid$DEM = 0
coordinates(data.grid) <- ~ x + y
crs(data.grid) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
data.grid
```



# crop DEM topography raster to the same extent as the river and measured water levels 
```{r}
# create a empty raster to store topography DEM 
data.raster <- rasterFromXYZ(data.frame(data.grid@coords), res=c(500,500))
dim(data.raster) = c(992,820)
extent(data.raster) = c(5749040.9338320437818766, 6159040.9338320437818766,19251409.7630450688302517, 19747409.7630450688302517)
data.raster[] = 0
data.raster

DEM_raster <- raster(ncols = 1000, nrows= 1000)
crs(DEM_raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
extent(DEM_raster) <- extent(data.grid)
DEM.raster <- resample(DEM, DEM_raster, method = "bilinear")
DEM.raster
    
```

# Data wrangling, select specific cols for subsequent analysis and convert to spatial point data frame with correct CRS
```{r}
library(dplyr)
library(pracma)
# choose the year of WL you want to work on
data <- water_level %>%
  dplyr::select(Longitude, Latitude, SWN, DEM, Depth, ScrTop1, ScrBot1, ScrTop2, ScrBot2, SH_Note, Aq2Use, WL2019,WL2020, Distance, Stream_WL) %>%
  # filter out NAs in both WL2019 and WL2020 to keep same number of measured water levels for both years
  filter(WL2019 != 0 & WL2020 != 0) 
head(data)

# we can also only exclude NAs for individual years
data_2019 <- data %>% 
  dplyr::select(Longitude, Latitude, WL2019) %>%
  filter(WL2019 != 'NA')

data_2020 <- data %>% 
  dplyr::select(Longitude, Latitude, WL2020) %>%
  filter(WL2020 != 'NA')



# create spatial point data frame with geographic coordinate systems
data_sp2019 <- SpatialPointsDataFrame(SpatialPoints(data_2019[,1:2]), data.frame(data_2019[,3]))
crs(data_sp2019) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

data_sp2020 <- SpatialPointsDataFrame(SpatialPoints(data_2020[,1:2]), data.frame(data_2020[,3]))
crs(data_sp2020) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

# create spatial point data frames with projected coordinate systems
new_crs <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
data_sp2019 <- spTransform(data_sp2019,new_crs)
data_sp2020 <- spTransform(data_sp2020,new_crs)

# rename the colnames
colnames(data_sp2019@coords)[1:2] = c("x","y")
colnames(data_sp2019@data)        = c("WL2019")  
colnames(data_sp2020@coords)[1:2] = c("x","y")
colnames(data_sp2020@data)        = c("WL2020")  

# create combined spatial point data frame for both years (2019 & 2020) with projected coordinate systems
data_spadf <- SpatialPointsDataFrame(SpatialPoints(data[,1:2]),data=data[,12:15])
crs(data_spadf) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
data_spadf <- spTransform(data_spadf,new_crs)
data_spadf@bbox <- t(matrix(extent(data.grid),2))  #crop
data_spadf


#Normalized Data
normalized = function (x) {(x-min(x))/(max(x)-min(x))}                              # create a normalization function
data_sp2019@data$WL2019n <- normalized(data_sp2019@data$WL2019)
data_sp2020@data$WL2020n <- normalized(data_sp2020@data$WL2020)
data_spadf@data$WL2019n  <- normalized(data_spadf@data$WL2019)
data_spadf@data$WL2020n  <- normalized(data_spadf@data$WL2020)

data_regdf = as.data.frame(data_spadf)
head(data_regdf)
```


# Display histogram of the covariates (DEM) and target variables (measured water level)
```{r}
# convert DEM raster to spatialpointdataframe
DEM.data <- rasterToPoints(DEM.raster, spatial=TRUE, progress="window")
DEM.data@bbox <- t(matrix(extent(data.grid),2))
# change the colname for the spatialpointdataframe
colnames(DEM.data@data)[1] = "DEM"
# normalized the data
DEM.data@data$DEMn <- normalized(DEM.data@data$DEM)
DEM.data

# convert to a dataframe
DEM.dataframe <- as.data.frame(DEM.data)
colnames(DEM.dataframe) <- c("DEM","DEMn", "x", "y")
# select 1 in every 1000 data for variogram modeling 
DEM.dataframesubset <- subset(DEM.dataframe[seq(1,nrow(DEM.dataframe), 1000),])
str(DEM.dataframesubset)
head(DEM.dataframesubset)

# create a subset DEM spatialpointdataframe for variogram modeling 
DEM.datasubset <- SpatialPointsDataFrame(SpatialPoints(DEM.dataframesubset[,3:4]),data=data.frame(DEM.dataframesubset[,1:2]))
DEM.datasubset@bbox <- t(matrix(extent(data.grid),2))
crs(DEM.datasubset) <- crs(DEM.data)
DEM.datasubset

plot(DEM.raster)
# plot the histograms to compare distributions 
h1 <- histogram(~ DEM, DEM.dataframe, xlab="DEM", col="lightblue",main ="DEM", nint=12)
#h2 <- histogram(~ DEM, river.data, xlab="River_DEM", col="red4", main = "River DEM", nint=12)
breaks = seq(150, 550, 25)
h3 <- histogram(~ WL2020, data_regdf, xlab="Measured Water Level (2020)", main = "Measured Water Level", col="green", breaks = breaks)
h4 <- histogram(~ WL2019, data_regdf, xlab="Measured Water Level (2019)", main = "Measured Water Level", col="green", breaks = breaks)
length(data_regdf$WL2020)
h1
#h2
h3
h4

```


## Co-kriging with topography (DEM) 
```{r}
plot(v.DEMn <- variogram(DEMn ~ 1, DEM.datasubset, cutoff=300000, width=10000), main = "Empirical Variogram w/ DEMn", pl = T)

# model by eye
m.DEMn <- vgm(.03, "Gau", 250000, .001)
# fit
(m.DEMn.f <- fit.variogram(v.DEMn, m.DEMn))

plot(v.DEMn, pl=T, model=m.DEMn.f, main = "Fitted Topography DEM Variogram")
```

```{r}
# create the cross-variogram for measured water level and topography
(g <- gstat(NULL, id = "WL", form = WL2019n ~ 1, data=data_spadf))

(g <- gstat(g, id = "TOPO", form = DEMn ~ 1, data=DEM.datasubset))

# Compute and display the two direct variogram and one cross-variogram
v.cross <- variogram(g)
str(v.cross)
plot(v.cross, pl=T)
```

```{r}
(g <- gstat(g, id = "WL_DEM", model = m.DEMn.f, fill.all=T))
# To ensure a positive definite co-krigingg system, use fit.lmc (fit linear model of co-regionalization)
# This takes the initial estimate, fits all the variograms, and then each of the partial sills is adjusted 
(g <- fit.lmc(v.cross, g, fit.method=1, correct.diagonal=1.01))
g
plot(variogram(g), model=g$model)


```




```{r}
k.cc <- predict(g, data.grid)
k.cc@data$WL.pred <- ifelse(k.cc@data$WL.pred >= 1.0, 1, k.cc@data$WL.pred)
k.cc@data$WL.pred <- ifelse(k.cc@data$WL.pred <= 0.0, 0, k.cc@data$WL.pred)
k.cc@data$WLn.pred = InvNormalize(k.cc@data$WL.pred)
k.cc
ko_pred <- data.frame(data = k.cc$WLn.pred, k.cc@coords)
ko_pred

sp::spplot(k.cc, "WLn.pred", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Co-Kirging Estimates with DEM", xlab = "x", ylab = "y", widths = 10)
```
# extract values for each acquifer for cokriging with DEM
```{r}
# extract values from spatial polygons
Yegua_Jackson_df      <- as.data.frame(k.cc[Yegua_Jackson,])
Sparta_df             <- as.data.frame(k.cc[Sparta,])
Queen_City_df         <- as.data.frame(k.cc[Queen_City,])
Carrizo_df           <- as.data.frame(k.cc[Carrizo,])
Calvert_Bluff_df      <- as.data.frame(k.cc[Calvert_Bluff,])
Simboro_df        <- as.data.frame(k.cc[Simboro,])
Hopper_df             <- as.data.frame(k.cc[Hopper,])
ShallowZone_df        <- as.data.frame(k.cc[ShallowZone,]) 
# set the min and max of measured water levels with correct year
min <- min(data_spadf@data$WL2019)
max <- max(data_spadf@data$WL2019)

# back transform to original scales
InvNormalize <- function (x) {x * (max - min) + min}

# find one average value for each acquifer
tol_avg   <- mean(InvNormalize(k.cc$WL.pred))
YJ_avg    <- mean(InvNormalize(Yegua_Jackson_df$WL.pred))
SPA_avg   <- mean(InvNormalize(Sparta_df$WL.pred))
QC_avg    <- mean(InvNormalize(Queen_City_df$WL.pred))
CAR_avg   <- mean(InvNormalize(Carrizo_df$WL.pred))
CB_avg    <- mean(InvNormalize(Calvert_Bluff_df$WL.pred))
SIM_avg   <- mean(InvNormalize(Simboro_df$WL.pred))


HOP_avg   <- mean(InvNormalize(Hopper_df$WL.pred))
SHAZ_avg  <- mean(InvNormalize(ShallowZone_df$WL.pred))
CO_avg <- data.frame(name = c("YJ", "SPA", "QC", "CAR", "CB", "SIM", "HOP", "TOL", "SHAZ"), avg = c(YJ_avg, SPA_avg, QC_avg, CAR_avg, CB_avg, SIM_avg, HOP_avg, tol_avg, SHAZ_avg))
CO_avg

```

# create raster and write on file
```{r}
# create raster and store it 
k.cc
k.temp <- k.cc[,-(1:5)]
k.temp 
# ck.raster <- raster(nrows = 10000,ncols =10000)
# crs(ck.raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
# extent(ck.raster) <- extent(ko)
# ck.raster <- rasterize(k.c, ck.raster, field = "WL2020.predN", fun = mean)
# ck.raster
#writeRaster(ck.raster, "CKDEM_raster_2020", "GTiff", overwrite = TRUE)
image_path = "C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Output Shapefile"
writeOGR(obj = k.temp, layer = "CK_TOPO_2019", dsn = image_path, driver = "ESRI Shapefile", overwrite_layer = TRUE)

```


















































