---
title: "Co-Kriging with Topography and Perennial Streams"
author: "YX"
date: "8/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(rgdal)
library(smerc)                                 # Implements statistical methods for analyzing the counts of areal data, with a 
library(maptools)
library(sf)
library(raster)
library(rasterVis)                             # The rasterVis package complements the raster package, providing a set of methods for enhanced     library(dismo)
library(spatial)
library(sp)
library(maxstat)
library(rgdal)
library(RColorBrewer)
library(LaplacesDemon)
library(imager)
library(pracma)
library(maps)
library(viridis)
library("rnaturalearth")
library("rnaturalearthdata")
library(ggspatial)
library(ggplot2)
library(wesanderson)
library(mltools)
library(RColorBrewer)
library(gganimate)                             # create anamination
library(tools)
library(maps)
library(zoo)
library(lubridate)
library(ggnewscale)                            # use multiple colour and fill scales in ggplot2
library(gstat)
library(taRifx)
library(splines)

## Read in the water level .csv file
path_waterlevel = ("input/data/water_levels")
water_level     <- read.csv(paste(path_waterlevel,"/Master_WellInfo_w_ShallowWLdata_1Yr_wBRAA_v2.csv",sep=''))
RiverRegression <- read.csv(paste(path_waterlevel,"/Linear_Relationship_of_River_Elevation.csv", sep = ''))

## read in the spatial polygons for each acquifer
path <- ("input/GIS/shapefiles/geology/District_Geology_Boundaries")
Yegua_Jackson <- readOGR(dsn = path, layer = "YJ_400_original")
Sparta        <- readOGR(dsn = path, layer = "SP_400_original")
Queen_City    <- readOGR(dsn = path, layer = "QC_400_original")
Carrizo       <- readOGR(dsn = path, layer = "CZ_400_original")
Calvert_Bluff <- readOGR(dsn = path, layer = "CB_400_original")
Simboro       <- readOGR(dsn = path, layer = "SB_400_original")
Hopper        <- readOGR(dsn = path, layer = "HP_400_original")
ShallowZone   <- readOGR(dsn = path, layer = "Total_400_original")
crs(ShallowZone) <- crs(Hopper)

## read in the spatial line df for river
path_river = ("input/GIS/shapefiles/hydrology")
#river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_GAM")
river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_pts_clip")

## read the DEM .tif file
# cliped DEM .tif file
#DEM <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/rasters/500ft_res/DEMs/DEM_ShallowArea_Resample500.tif")
#DEM <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/rasters/DEM_POSGCD_NeighboringCounties.tif")
DEM <- raster("input/GIS/rasters/DEM_ShallowArea_Resample500.tif") # hpham: This script reads a different DEM? 


D2River_Grid = readOGR(dsn = path_river, layer = "GridCells_500_within_6000_ft_Smooth_Perennial_Streams_wNearestSmoothStreamPt")

# import the pumping raster (difference in water level after pumping )
pumping500_2018 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2018.500ft.pst15.tif")
pumping500_2019 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2019.500ft.pst15.tif")
pumping500_2020 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")

# import the GAM (simulated water levels) 
sim_WL2020 <- raster("input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2020.500ft.pst15.tif")
sim_WL2019 <- raster("input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2019.500ft.pst15.tif")
#plot(sim_WL2020)
#plot(sim_WL2019)



#setwd("c:/Users/hpham/OneDrive - INTERA Inc/projects/012_rscript_geostats_s/scripts/")


```


# create grids for kriging interpolations 
```{r}
# x <- seq(5800000, 6100000,400)
# y <- seq(19333000, 19633000,400)

x <- seq(5749040.9338320437818766, 6159040.9338320437818766,500)
y <- seq(19251409.7630450688302517, 19747409.7630450688302517,500)
data.grid <- meshgrid(x,y)

data.grid <- data.frame(unlist(matrix(data.grid$X)), unlist(matrix(data.grid$Y)))
colnames(data.grid) <- c("x","y")
data.grid$DEM = 0
coordinates(data.grid) <- ~ x + y
crs(data.grid) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
data.grid
```


# crop DEM topography raster to the same extent as the river and measured water levels 
```{r}
# create a empty raster to store topography DEM 
data.raster <- rasterFromXYZ(data.frame(data.grid@coords), res=c(500,500))
dim(data.raster) = c(992,820)
extent(data.raster) = c(5749040.9338320437818766, 6159040.9338320437818766,19251409.7630450688302517, 19747409.7630450688302517)
data.raster[] = 0
data.raster

DEM_raster <- raster(ncols = 1000, nrows= 1000)

crs(DEM_raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
extent(DEM_raster) <- extent(data.grid)
DEM.raster <- resample(DEM, DEM_raster, method = "bilinear")
DEM.raster
    
```


# Data wrangling, select specific cols for subsequent analysis and convert to spatial point data frame with correct CRS
```{r}
library(dplyr)
library(pracma)
# choose the year of WL you want to work on
data <- water_level %>%
  dplyr::select(Longitude, Latitude, SWN, DEM, Depth, ScrTop1, ScrBot1, ScrTop2, ScrBot2, SH_Note, Aq2Use, WL2019,WL2020) %>%
  # filter out NAs in both WL2019 and WL2020 to keep same number of measured water levels for both years
  filter(WL2019 != 0 & WL2020 != 0) 
head(data)

# we can also only exclude NAs for individual years
data_2019 <- data %>% 
  dplyr::select(Longitude, Latitude, WL2019) %>%
  filter(WL2019 != 'NA')

data_2020 <- data %>% 
  dplyr::select(Longitude, Latitude, WL2020) %>%
  filter(WL2020 != 'NA')



# create spatial point data frame with geographic coordinate systems
data_sp2019 <- SpatialPointsDataFrame(SpatialPoints(data_2019[,1:2]), data.frame(data_2019[,3]))
crs(data_sp2019) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

data_sp2020 <- SpatialPointsDataFrame(SpatialPoints(data_2020[,1:2]), data.frame(data_2020[,3]))
crs(data_sp2020) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

# create spatial point data frames with projected coordinate systems
new_crs <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
data_sp2019 <- spTransform(data_sp2019,new_crs)
data_sp2020 <- spTransform(data_sp2020,new_crs)

# rename the colnames
colnames(data_sp2019@coords)[1:2] = c("x","y")
colnames(data_sp2019@data)        = c("WL2019")  
colnames(data_sp2020@coords)[1:2] = c("x","y")
colnames(data_sp2020@data)        = c("WL2020")  

# create combined spatial point data frame for both years (2019 & 2020) with projected coordinate systems
data_spadf <- SpatialPointsDataFrame(SpatialPoints(data[,1:2]),data=data[,12:13])
crs(data_spadf) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
data_spadf <- spTransform(data_spadf,new_crs)
data_spadf@bbox <- t(matrix(extent(data.grid),2))  #crop
data_spadf

data_regdf = as.data.frame(data_spadf)
#Normalized Data
normalized = function (x) {(x-min(x))/(max(x)-min(x))}                              # create a normalization function
data_sp2019@data$WL2019n <- normalized(data_sp2019@data$WL2019)
data_sp2020@data$WL2020n <- normalized(data_sp2020@data$WL2020)
data_spadf@data$WL2019n  <- normalized(data_spadf@data$WL2019)
data_spadf@data$WL2020n  <- normalized(data_spadf@data$WL2020)
```

# Display histogram of the covariates (DEM) and target variables (measured water level)
```{r}
# convert DEM raster to spatialpointdataframe
DEM.data <- rasterToPoints(DEM.raster, spatial=TRUE, progress="window")
DEM.data@bbox <- t(matrix(extent(data.grid),2))
# change the colname for the spatialpointdataframe
colnames(DEM.data@data)[1] = "DEM"
# normalized the data
DEM.data@data$DEMn <- normalized(DEM.data@data$DEM)
DEM.data
# convert to a dataframe
DEM.dataframe <- as.data.frame(DEM.data)
colnames(DEM.dataframe) <- c("DEM","DEMn", "x", "y")
DEM.dataframesubset <- subset(DEM.dataframe[seq(1,nrow(DEM.dataframe), 800),])
str(DEM.dataframesubset)
head(DEM.dataframesubset)

# create a subset DEM spatialpointdataframe for variogram modeling 
DEM.datasubset <- SpatialPointsDataFrame(SpatialPoints(DEM.dataframesubset[,3:4]),data=data.frame(DEM.dataframesubset[,1:2]))
DEM.datasubset@bbox <- t(matrix(extent(data.grid),2))
crs(DEM.datasubset) <- crs(DEM.data)
DEM.datasubset

plot(DEM.raster)
# plot the histograms to compare distributions 
h1 <- histogram(~ DEM, DEM.dataframe, xlab="DEM", col="lightblue",main ="DEM", nint=12)

h3 <- histogram(~ WL2020, data_regdf, xlab="Measured Water Level (2020)", main = "Measured Water Level", col="green", nint=12)
h1

h3
```

# Modelling the spatial structure of the target variable - make variogram
# 1. the variogram method computes the experimental variogram as a variogram object
# 2. the vgm method creates a variogram model object
# 3. the fit.variogram method adjusts a variogram model object to a variogram object
```{r}
# look at the spatial point data frame
data_spadf$WL2020n

# plot the empirical variogram for measured water levels at specific years
plot(v.WLn <- variogram(WL2019n ~ 1 , data=data_spadf, cutoff=300000, width=10000), main = "Empirical Variogram",pl = T)
# create a model for the empirical variogram
m.WLn <- vgm(0.05,"Gau",200000,0.003)
# plot both the fitted variogram model with the empirical variogram
plot(v.WLn, pl=T, model=m.WLn, main = "Variogram Model")
(m.WLn.f <- fit.variogram(v.WLn, m.WLn))

plot(v.WLn, pl=T, model=m.WLn.f, main = "Fitted OK Variogram Model")
```

## Ordinary Kriging Interpolation of the Target Variable
```{r}
# use krige function for oridinary kriging 
ko <- krige(WL2019n~1, locations=data_spadf, newdata=data.grid, model=m.WLn.f)
# extract the kriging estimates and kriging variances 
ko_pred <- data.frame(ko$var1.pred)
ko_var <- data.frame(ko$var1.var)

# find the min and max of the original scale measured water levels
min <- min(data_spadf@data$WL2019)
max <- max(data_spadf@data$WL2019)
# back transform to original scales
InvNormalize <- function (x) {x * (max - min) + min}

ko$var1.predn = InvNormalize(ko$var1.pred)

# use spplot to plot the kriging results 
sp::spplot(ko, "var1.pred", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Ordinary Kirging Estimates")
sp::spplot(ko, "var1.var", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Ordinary Kirging Estimates")

```

# plot contours
```{r}
ko.dataframe = as.data.frame(ko)
pred <- ggplot(ko.dataframe, aes(x ,y, z = var1.pred)) +
  geom_contour_filled(bins = 20)
var <- ggplot(ko.dataframe, aes(x ,y, z = var1.var)) +
  geom_contour_filled(bins = 20)
pred
```

# create a spline model that model the distance relationship of river elevation 
```{r}
# filter out 5000 ft
RiverRegression_subset = RiverRegression %>%
  dplyr::filter(Distance <= 5000)

# set the values to numeric
RiverRegression_subset$median = as.numeric(RiverRegression_subset$median)

# create a spline model
summary(model1 <- glm(median ~ bs(Distance, df = 3), data = RiverRegression_subset, family = "gaussian"))

# using the spline model to make prediction
Predicted_RiverElevation = predict(model1, data.frame(Distance = RiverRegression_subset$Distance))

RiverRegression_subset$prediction = Predicted_RiverElevation
# plot the prediction and the data
ggplot(data = RiverRegression_subset) +
  geom_point(aes(x = Distance, y = median)) +
  geom_line(aes(x = Distance, y = prediction))


```

```{r}
# convert spatial point data frame into simple feature in sf package
ko_sf = st_as_sf(ko)
D2River_Grid_sf = st_as_sf(D2River_Grid)


ko_sfc = st_as_sfc(ko)
D2River_Grid_sfc = st_as_sfc(D2River_Grid)


# create a buffer zone
ko_sfc_buf = st_buffer(ko_sfc, 2000)



# find the intersection of kriging data frame and river data frame 
# create a sparse geometry binary predicate, also called sparse matrix, which is a list with integer vectors only holding the indices for each polygon that intersects. 
# find x that s in y
koRiver_sgbp = st_intersects(x = ko_sfc_buf, y= D2River_Grid_sfc)
Riverko_sgbp = st_intersects(x = D2River_Grid_sfc, y = ko_sfc_buf)

class(koRiver_sgbp)

koRiver_logical = lengths(koRiver_sgbp) > 0
River_logical = lengths(Riverko_sgbp) > 0 

ko_river_intersect = ko_sf[koRiver_logical,]
river_ko_intersect = D2River_Grid_sf[River_logical,]

ko_river_intersect
river_ko_intersect 


sum(is.na(river_ko_intersect$SmoothWL))

#ko_river_intersect$river_pred = 0


#river_ko_intersect[is.na(river_ko_intersect$SmoothWL),]$SmoothWL = rep(mean(D2River_Grid_sf$SmoothWL),length(river_ko_intersect[is.na(river_ko_intersect$SmoothWL),]$SmoothWL))

ko_river_intersect$river_pred = (river_ko_intersect$SmoothWL +  predict(model1, data.frame(Distance = river_ko_intersect$Distance))) * 0.8 + ko_river_intersect$var1.predn * 0.2

sum(is.na(ko_river_intersect$river_pred))

diff = ko_river_intersect$river_pred - ko_river_intersect$var1.predn
hist(river_ko_intersect$SmoothWL)




ko_spatial = as_Spatial(ko_river_intersect)
ko.dataframe = as.data.frame(ko_spatial)
pred <- ggplot(ko.dataframe, aes(coords.x1 ,coords.x2, z = river_pred)) +
  geom_contour_filled(bins = 20)
pred
D2River_dataframe = as.data.frame(D2River_Grid)
pred <- ggplot(D2River_dataframe, aes(coords.x1 ,coords.x2, z = SmoothWL)) +
  geom_contour_filled(bins = 20)
pred

# normalize the column
normalized = function (x) {(x-min(x))/(max(x)-min(x))}    
ko_spatial$river_predn = normalized(ko_spatial@data$river_pred)
plot(ko_river_intersect)


```

```{r}
# convert to a dataframe
ko.dataframe <- as.data.frame(ko_spatial)
#colnames(DEM.dataframe) <- c("DEM","DEMn", "x", "y")
# select 1 in every 1000 data for variogram modeling 
set.seed(123)
ko.dataframesubset <- subset(ko.dataframe[seq(1,nrow(ko.dataframe), 300),])

str(ko.dataframesubset)

# create a subset DEM spatialpointdataframe for variogram modeling 
ko.datasubset <- SpatialPointsDataFrame(SpatialPoints(ko.dataframesubset[,6:7]),data=data.frame(ko.dataframesubset[,4:5]))
ko.datasubset@bbox <- t(matrix(extent(data.grid),2))
crs(ko.datasubset) <- crs(DEM.data)
ko.datasubset
```

## Co-kriging with topography (DEM) 
```{r}
plot(v.DEMn <- variogram(DEMn ~ 1, DEM.datasubset, cutoff=300000, width=20000), main = "Empirical Variogram w/ DEMn", pl = T)

# model by eye
m.DEMn <- vgm(.025, "Gau", 250000, .001)
# fit
(m.DEMn.f <- fit.variogram(v.DEMn, m.DEMn))

plot(v.DEMn, pl=T, model=m.DEMn.f, main = "Fitted Topography DEM Variogram")
```


```{r}
(g <- gstat(NULL, id = "WL.RIVER", form = river_predn ~ 1, data=ko.datasubset))

(g <- gstat(g, id = "DEM", form = DEMn ~ 1, data=DEM.datasubset))

# Compute and display the two direct variogram and one cross-variogram

v.cross <- variogram(g)
str(v.cross)
plot(v.cross, pl=T)

```

```{r}
(g <- gstat(g, id = "WL.RIVER", model = m.DEMn.f, fill.all=T))
# To ensure a positive definite co-krigingg system, use fit.lmc (fit linear model of co-regionalization)
# This takes the initial estimate, fits all the variograms, and then each of the partial sills is adjusted 
(g <- fit.lmc(v.cross, g, fit.method=1, correct.diagonal=1.01))
g
plot(variogram(g), model=g$model)


```




```{r}
k.cc <- predict(g, data.grid)
k.cc@data$WL.RIVER.pred <- ifelse(k.cc@data$WL.RIVER.pred >= 1.0, 1, k.cc@data$WL.RIVER.pred)
k.cc@data$WL.RIVER.pred <- ifelse(k.cc@data$WL.RIVER.pred <= 0.0, 0, k.cc@data$WL.RIVER.pred)
k.cc@data$WL.RIVER.predn = InvNormalize(k.cc@data$WL.RIVER.pred)
k.cc


sp::spplot(k.cc, "WL.RIVER.pred", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Co-Kirging Estimates with DEM", xlab = "x", ylab = "y", widths = 10)
```




# extract values for each acquifer for cokriging with DEM
```{r}
# extract values from spatial polygons
Yegua_Jackson_df      <- as.data.frame(k.cc[Yegua_Jackson,])
Sparta_df             <- as.data.frame(k.cc[Sparta,])
Queen_City_df         <- as.data.frame(k.cc[Queen_City,])
Carrizo_df           <- as.data.frame(k.cc[Carrizo,])
Calvert_Bluff_df      <- as.data.frame(k.cc[Calvert_Bluff,])
Simboro_df        <- as.data.frame(k.cc[Simboro,])
Hopper_df             <- as.data.frame(k.cc[Hopper,])
ShallowZone_df        <- as.data.frame(k.cc[ShallowZone,]) 
min <- min(data_spadf@data$WL2020)
max <- max(data_spadf@data$WL2020)
# back transform to original scales
InvNormalize <- function (x) {x * (max - min) + min}

# find one average value for each acquifer
tol_avg   <- mean(k.cc$WL.RIVER.predn )
YJ_avg    <- mean(Yegua_Jackson_df$WL.RIVER.predn )
SPA_avg   <- mean(Sparta_df$WL.RIVER.predn )
QC_avg    <- mean(Queen_City_df$WL.RIVER.predn )
CAR_avg   <- mean(Carrizo_df$WL.RIVER.predn )
CB_avg    <- mean(Calvert_Bluff_df$WL.RIVER.predn )
SIM_avg   <- mean(Simboro_df$WL.RIVER.predn )
HOP_avg   <- mean(Hopper_df$WL.RIVER.predn )
SHAZ_avg  <- mean(ShallowZone_df$WL.RIVER.predn )
CO_avg <- data.frame(name = c("YJ", "SPA", "QC", "CAR", "CB", "SIM", "HOP", "TOL", "SHAZ"), avg = c(YJ_avg, SPA_avg, QC_avg, CAR_avg, CB_avg, SIM_avg, HOP_avg, tol_avg, SHAZ_avg))
CO_avg

```

# create raster and write on file
```{r}
# create raster and store it 
k.cc@data$WL2020.predN <- InvNormalize(k.cc$WL2020.pred)
k.cc
k.temp <- k.cc[,-(1:5)]
k.temp
# ck.raster <- raster(nrows = 10000,ncols =10000)
# crs(ck.raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
# extent(ck.raster) <- extent(ko)
# ck.raster <- rasterize(k.c, ck.raster, field = "WL2020.predN", fun = mean)
# ck.raster
#writeRaster(ck.raster, "CKDEM_raster_2020", "GTiff", overwrite = TRUE)
image_path = "C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Output Shapefile"
writeOGR(obj = k.temp, layer = "CK_TOPORIVER_2019", dsn = image_path, driver = "ESRI Shapefile", overwrite_layer = TRUE)

```






























