---
title: "Co-Kriging with TOPO RIVER PUMPING"
author: "YX"
date: "8/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rgdal)
library(smerc)                                 # Implements statistical methods for analyzing the counts of areal data, with a 
library(maptools)
library(sf)
library(raster)
library(rasterVis)                             # The rasterVis package complements the raster package, providing a set of methods for enhanced     library(dismo)
library(spatial)
library(sp)
library(maxstat)
library(rgdal)
library(RColorBrewer)
library(LaplacesDemon)
library(imager)
library(pracma)
library(maps)
library(viridis)
library("rnaturalearth")
library("rnaturalearthdata")
library(ggspatial)
library(ggplot2)
library(wesanderson)
library(mltools)
library(RColorBrewer)
library(gganimate)                             # create anamination
library(tools)
library(maps)
library(zoo)
library(lubridate)
library(ggnewscale)                            # use multiple colour and fill scales in ggplot2
library(gstat)
library(taRifx)
library(splines)

## Read in the water level .csv file
#path_waterlevel = ("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data")
#water_level     <- read.csv(paste(path_waterlevel,"/Master_WellInfo_w_ShallowWLdata_1Yr_wBRAA_v2.csv",sep=''))
GAM_sim_WL_yr <- as.integer(2020)
data_col_name <- "WL2020"
thre_cutoff <- as.integer(300000)
thre_width <- as.integer(10000)

path_waterlevel = ("input/data/water_levels")
ifile_WL <- paste(path_waterlevel,"/Shallow.Wells.GMA12.YJ.hds.Smooth_",GAM_sim_WL_yr,"_hp.csv",sep='')
water_level <- read.csv(ifile_WL)


RiverRegression <- read.csv(paste(path_waterlevel,"/Linear_Relationship_of_River_Elevation.csv", sep = ''))

# ## read in the spatial polygons for each acquifer
# path <- ("input/GIS/shapefiles/geology/District_Geology_Boundaries")
# Yegua_Jackson <- readOGR(dsn = path, layer = "YJ_400_original")
# Sparta        <- readOGR(dsn = path, layer = "SP_400_original")
# Queen_City    <- readOGR(dsn = path, layer = "QC_400_original")
# Carrizo       <- readOGR(dsn = path, layer = "CZ_400_original")
# Calvert_Bluff <- readOGR(dsn = path, layer = "CB_400_original")
# Simboro       <- readOGR(dsn = path, layer = "SB_400_original")
# Hopper        <- readOGR(dsn = path, layer = "HP_400_original")
# ShallowZone   <- readOGR(dsn = path, layer = "Total_400_original")
# crs(ShallowZone) <- crs(Hopper)

# [3] Read in the spatial line df for river -----------------------------------
path_river = ("input/GIS/shapefiles/hydrology")
#river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_GAM")
river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_pts_clip")

## read the DEM .tif file
# cliped DEM .tif file
DEM <- raster("input/GIS/rasters/DEM_ShallowArea_Resample500.tif") # hpham: This script reads a different DEM? #DEM <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/rasters/DEM_POSGCD_NeighboringCounties.tif")


D2River_Grid = readOGR(dsn = path_river, 
layer = "GridCells_500_within_6000_ft_Smooth_Perennial_Streams_wNearestSmoothStreamPt")

D2River_Grid2 = readOGR(dsn = path_river, 
layer = "GridCell_Corners_500_within_6000_ft_Smooth_Perennial_Streams_wNearestSmoothStreamPt")

#sp::spplot(D2River_Grid, "SmoothWL", pch=19, col = "transparent",
#           scales=list(draw=TRUE), main="SmoothWL")

# import the pumping raster (difference in water level after pumping )
#pumping500_2018 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2018.500ft.pst15.tif")
#pumping500_2019 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2019.500ft.pst15.tif")

#if (GAM_res == 500) {
  pumping500_2020 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")
  # Ross updated these raster on 08/26/2020
  ifile_GAM_WL <- "input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst18.YJ/Shallow.Combined.Hds.2020.500ft.pst18.YJ.tif"
  #sim_WL2020 <- raster("input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst18.YJ/Shallow.Combined.Hds.2020.500ft.pst18.YJ.tif")
#} else if (GAM_res == 32000000000000) {
#  pumping500_2020 <- raster("input/GIS/rasters/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")
#  # last updated: 09/04/2020 by Ross
#  #ifile_GAM_WL = "input/GIS/Smoothing.GAM.Rasters/Shallow.Below.Hds.2020.Smooth32000ft.tif"
#  ifile_GAM_WL = "input/GIS/Smoothing.GAM.Rasters/Shallow.Combine.Hds.2020.Smooth32000ft.tif"
#  
#}
# import the GAM (simulated water levels) 
#sim_WL2020 <- raster("input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2020.500ft.pst15.tif")
#sim_WL2019 <- raster("input/GIS/rasters/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2019.500ft.pst15.tif")

sim_WL2020 <- raster(ifile_GAM_WL)
#plot(sim_WL2020)
#plot(sim_WL2019)

# import pumping drawdown information 
pumpingdrawdown_path = ("input/data/pumping")
#pumping_drawdown2018 = readOGR(dsn = pumpingdrawdown_path, layer = "Pts.Shallow.Below.DD.2018.500ft.pst18.YJ")
#pumping_drawdown2019 = readOGR(dsn = pumpingdrawdown_path, layer = "Pts.Shallow.Below.DD.2019.500ft.pst18.YJ")
pumping_drawdown2020 = readOGR(dsn = pumpingdrawdown_path, layer = "Pts.Shallow.Below.DD.2020.500ft.pst18.YJ")



```


# create grids for kriging interpolations 
```{r}
# x <- seq(5800000, 6100000,400)
# y <- seq(19333000, 19633000,400)

x <- seq(5749040.9338320437818766, 6159040.9338320437818766,500)
y <- seq(19251409.7630450688302517, 19747409.7630450688302517,500)
data.grid <- meshgrid(x,y)

data.grid <- data.frame(unlist(matrix(data.grid$X)), unlist(matrix(data.grid$Y)))
colnames(data.grid) <- c("x","y")
data.grid$DEM = 0
coordinates(data.grid) <- ~ x + y
crs(data.grid) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
data.grid
```


# crop DEM topography raster to the same extent as the river and measured water levels 
```{r}
# create a empty raster to store topography DEM 
data.raster <- rasterFromXYZ(data.frame(data.grid@coords), res=c(500,500))
dim(data.raster) = c(992,820)
extent(data.raster) = c(5749040.9338320437818766, 6159040.9338320437818766,19251409.7630450688302517, 19747409.7630450688302517)
data.raster[] = 0
data.raster

DEM_raster <- raster(ncols = 1000, nrows= 1000)

crs(DEM_raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
extent(DEM_raster) <- extent(data.grid)
DEM.raster <- resample(DEM, DEM_raster, method = "bilinear")
DEM.raster
    
```


# Data wrangling, select specific cols for subsequent analysis and convert to spatial point data frame with correct CRS
```{r}
library(dplyr)
library(pracma)
# choose the year of WL you want to work on
#data <- water_level %>%
#  dplyr::select(Longitude, Latitude, SWN, DEM, Depth, ScrTop1, ScrBot1, ScrTop2, ScrBot2, SH_Note, Aq2Use, WL2019,WL2020) %>%
  # filter out NAs in both WL2019 and WL2020 to keep same number of measured water levels for both years
#  filter(WL2019 != 0 & WL2020 != 0) 
#head(data)

data <- water_level %>%
  dplyr::select(Longitude, Latitude, data_col_name) # hpham


# we can also only exclude NAs for individual years
#data_2019 <- data %>% 
#  dplyr::select(Longitude, Latitude, WL2019) %>%
#  filter(WL2019 != 'NA')

data_2020 <- data %>% 
  dplyr::select(Longitude, Latitude, WL2020) %>%
  filter(WL2020 != 'NA')



# create spatial point data frame with geographic coordinate systems
#data_sp2019 <- SpatialPointsDataFrame(SpatialPoints(data_2019[,1:2]), data.frame(data_2019[,3]))
#crs(data_sp2019) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

data_sp2020 <- SpatialPointsDataFrame(SpatialPoints(data_2020[,1:2]), data.frame(data_2020[,3]))
crs(data_sp2020) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

# create spatial point data frames with projected coordinate systems
new_crs <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
#data_sp2019 <- spTransform(data_sp2019,new_crs)
data_sp2020 <- spTransform(data_sp2020,new_crs)

# rename the colnames
#colnames(data_sp2019@coords)[1:2] = c("x","y")
#colnames(data_sp2019@data)        = c("WL2019")  
colnames(data_sp2020@coords)[1:2] = c("x","y")
colnames(data_sp2020@data)        = c("WL2020")  

# create combined spatial point data frame for both years (2019 & 2020) with projected coordinate systems
#data_spadf <- SpatialPointsDataFrame(SpatialPoints(data[,1:2]),data=data[,12:13])
#crs(data_spadf) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
#data_spadf <- spTransform(data_spadf,new_crs)
#data_spadf@bbox <- t(matrix(extent(data.grid),2))  #crop
#data_spadf

#data_regdf = as.data.frame(data_spadf)
#Normalized Data
normalized = function (x) {(x-min(x))/(max(x)-min(x))}                              # create a normalization function
#data_sp2019@data$WL2019n <- normalized(data_sp2019@data$WL2019)
data_sp2020@data$WL2020n <- normalized(data_sp2020@data$WL2020)
#data_spadf@data$WL2019n  <- normalized(data_spadf@data$WL2019)
#data_spadf@data$WL2020n  <- normalized(data_spadf@data$WL2020)
```

# Display histogram of the covariates (DEM) and target variables (measured water level)
```{r}
# convert DEM raster to spatialpointdataframe
DEM.data <- rasterToPoints(DEM.raster, spatial=TRUE, progress="window")
DEM.data@bbox <- t(matrix(extent(data.grid),2))
# change the colname for the spatialpointdataframe
colnames(DEM.data@data)[1] = "DEM"
# normalized the data
DEM.data@data$DEMn <- normalized(DEM.data@data$DEM)
DEM.data
# convert to a dataframe
DEM.dataframe <- as.data.frame(DEM.data)
colnames(DEM.dataframe) <- c("DEM","DEMn", "x", "y")
DEM.dataframesubset <- subset(DEM.dataframe[seq(1,nrow(DEM.dataframe), 100),])
str(DEM.dataframesubset)
head(DEM.dataframesubset)

# create a subset DEM spatialpointdataframe for variogram modeling 
DEM.datasubset <- SpatialPointsDataFrame(SpatialPoints(DEM.dataframesubset[,3:4]),data=data.frame(DEM.dataframesubset[,1:2]))
DEM.datasubset@bbox <- t(matrix(extent(data.grid),2))
crs(DEM.datasubset) <- crs(DEM.data)
DEM.datasubset

plot(DEM.raster)

# plot the histograms to compare distributions 
h1 <- histogram(~ DEM, DEM.dataframe, xlab="DEM", col="lightblue",main ="DEM", nint=12)
h1

#h3 <- histogram(~ WL2020, data_sp2020, xlab="Measured Water Level (2020)", main = "Measured Water Level", col="green", nint=12)
#h3

base::save.image('cur.RData')


```


# Modelling the spatial structure of the target variable - make variogram
# 1. the variogram method computes the experimental variogram as a variogram object
# 2. the vgm method creates a variogram model object
# 3. the fit.variogram method adjusts a variogram model object to a variogram object
```{r}

#base::load('cur.RData')
v.WLn <- variogram(WL2020n ~ 1 , data=data_sp2020, cutoff=thre_cutoff, width=thre_width)
plot(v.WLn, main = "Empirical Variogram",pl = T)

# create a model for the empirical variogram
m.WLn <- vgm(0.05,"Gau",200000,0.003)

# plot both the fitted variogram model with the empirical variogram
plot(v.WLn, pl=T, model=m.WLn, main = "Variogram Model")
(m.WLn.f <- fit.variogram(v.WLn, m.WLn))

#plot(v.WLn, pl=T, model=m.WLn.f, main = "Fitted OK Variogram Model")

```

## Ordinary Kriging Interpolation of the Target Variable
```{r}
# use krige function for oridinary kriging 
ko <- krige(WL2020n~1, locations=data_sp2020, newdata=data.grid, model=m.WLn.f)

# extract the kriging estimates and kriging variances 
ko_pred <- data.frame(ko$var1.pred)
ko_var <- data.frame(ko$var1.var)

# find the min and max of the original scale measured water levels for each year
min <- min(data_sp2020@data$WL2020)
max <- max(data_sp2020@data$WL2020)

# back transform to original scales
InvNormalize <- function (x) {x * (max - min) + min}

ko$var1.predn = InvNormalize(ko$var1.pred)

# use spplot to plot the kriging results 
sp::spplot(ko, "var1.pred", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Ordinary Kirging Estimates")
sp::spplot(ko, "var1.var", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Ordinary Kirging Estimates")

```

# create a spline model that model the distance relationship of river elevation 
```{r}
# filter out distance longer than 5000 ft
RiverRegression_subset = RiverRegression %>%
  dplyr::filter(Distance <= 6000)

# set the values to numeric
RiverRegression_subset$median = as.numeric(RiverRegression_subset$median)

# create a spline model
summary(model1 <- glm(median ~ bs(Distance, df = 4), data = RiverRegression_subset, family = "gaussian"))

# using the spline model to make prediction
Predicted_RiverElevation = predict(model1, data.frame(Distance = RiverRegression_subset$Distance))

RiverRegression_subset$prediction = Predicted_RiverElevation
# plot the prediction and the data
ggplot(data = RiverRegression_subset) +
  geom_point(aes(x = Distance, y = median)) +
  geom_line(aes(x = Distance, y = prediction)) +
  labs(title = "The Median Value of River Elevation Away From River Streams", x = "Distance From River Streams (ft)", y = "Median Value of River Elevations (ft)")


```


# add the river elevation on the the prdinary kriging estimates of measured water levels
```{r}
# convert spatial point data frame into simple feature in sf package
sf_ok_cells = st_as_sf(ko)

#ofile = paste("output/", "sf_ok_cells.csv",sep="")
#write.csv(sf_ok_cells, ofile)

#D2River_Grid_sf = st_as_sf(D2River_Grid)
sf_river_cells = st_as_sf(D2River_Grid2)

# create a linear ratio of distance to assign approporiate weights
LinearDistance_ratio = 1 - sf_river_cells$Distance/ max(sf_river_cells$Distance)





#WL_DIFF_by_DIST <- predict(model1, data.frame(Distance = sf_river_cells$Distance))
#size(WL_DIFF_by_DIST)

#OK_pred_WL <- sf_ok$var1.predn

# [FINAL] calculate the river pred from distance and smooth water level from 
cof <- 0.8*LinearDistance_ratio
#LinearDistance_ratio <- 1

# sf_ok$WLpumping_river <- (sf_river_cells$SmoothWL +  predict(model1, data.frame(Distance = sf_river_cells$Distance))) * cof +  sf_ok$var1.predn*(1-cof)

#sf_river_cells$WL_DIFF = predict(model1, data.frame(Distance = sf_river_cells$Distance))

#sf_river_cells <- sf_river_cells %>% mutate(GWL_near_river = (SmoothWL + WL_DIFF)*cof)

#sf_ok_cells <- sf_ok_cells %>% mutate(OK_pred = var1.predn*(1-cof))
#plot(sf_ok)

sf_ok_cells <- sf_ok_cells %>% mutate(Xcoor = round(st_coordinates(sf_ok_cells)[,1],0), 
                                                    Ycoor = round(st_coordinates(sf_ok_cells)[,2],0))
sf_river_cells <- sf_river_cells %>% mutate(Xcoor = round(x,0), 
                                                    Ycoor = round(y,0))

join_sf <- as.data.frame(as_Spatial(sf_river_cells)) %>% 
  left_join(as.data.frame(as_Spatial(sf_ok_cells)), by=c("Xcoor", "Ycoor"))
str(join_sf)


df_ok_riv <- join_sf %>%
  dplyr::select(x, y, SmoothWL, Distance, var1.predn)
str(df_ok_riv)

df_ok_riv$WL_DIFF = predict(model1, data.frame(Distance = df_ok_riv$Distance))
df_ok_riv <- df_ok_riv %>% mutate(GWL_near_riv = (SmoothWL + WL_DIFF)*cof)
df_ok_riv <- df_ok_riv %>% mutate(OK_RIV_pred = GWL_near_riv + var1.predn*(1-cof))
str(df_ok_riv)

#join_df_clean <- join_df %>% 
#  filter(!is.na(GWL_near_river)) 

#ko_predWL <- as_tibble(as_Spatial(sf_ok))
#ko_predWL <- as.data.frame(as_Spatial(sf_ok_cells))
#str(ko_predWL)

#river_predWL <- as_tibble(as_Spatial(sf_river_cells))
#river_predWL <- as.data.frame(as_Spatial(sf_river_cells))
#str(river_predWL)

#join_df <- ko_predWL %>% 
#  left_join(river_predWL, by=c("Xcoor", "Ycoor"))
#str(join_df)

pred <- ggplot(df_ok_riv, aes(x ,y, z = OK_RIV_pred)) +
  geom_contour_filled(bins = 10) + ggtitle("Adjusted OK water levels based on river stages (ft)")
pred

pred <- ggplot(df_ok_riv, aes(x ,y, z = var1.predn)) +
  geom_contour_filled(bins = 10) + ggtitle("OK predicted water level (ft)")
pred

pred <- ggplot(df_ok_riv, aes(x ,y, z = SmoothWL)) +
  geom_contour_filled(bins = 10) + ggtitle("SmoothWL")
pred


# Save to a csv file
ofile = paste("output/", "df_ok_riv.csv",sep="")
write.csv(df_ok_riv, ofile)



# normalize the column
#normalized = function (x) {(x-min(x))/(max(x)-min(x))}    
#ko_spatial$WLpumping_rivern = normalized(ko_spatial@data$WLpumping_river )



```
