---
title: "Ordinary Kring with Measured Water Levels"
author: "Yuchen Xiao"
date: "8/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rgdal)
library(smerc)                                 # Implements statistical methods for analyzing the counts of areal data, with a 
library(maptools)
library(sf)
library(raster)
library(rasterVis)                             # The rasterVis package complements the raster package, providing a set of methods for enhanced         
library(dismo)
library(spatial)
library(sp)
library(maxstat)
library(rgdal)
library(RColorBrewer)
library(LaplacesDemon)
library(imager)
library(pracma)
library(maps)
library(viridis)
library("rnaturalearth")
library("rnaturalearthdata")
library(ggspatial)
library(ggplot2)
library(wesanderson)
library(mltools)
library(RColorBrewer)
library(gganimate)                             # create anamination
library(tools)
library(maps)
library(zoo)
library(lubridate)
library(ggnewscale)                            # use multiple colour and fill scales in ggplot2
library(gstat)
library(taRifx)

## Read in the water level .csv file
path_waterlevel = ("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data")
water_level <- read.csv(paste(path_waterlevel,"/Well_info.csv",sep=''))

## read in the spatial polygons for each acquifer
path = ("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/geology/District_Geology_Boundaries")
Yegua_Jackson <- readOGR(dsn = path, layer = "YJ_400_original")
Sparta        <- readOGR(dsn = path, layer = "SP_400_original")
Queen_City    <- readOGR(dsn = path, layer = "QC_400_original")
Carrizo       <- readOGR(dsn = path, layer = "CZ_400_original")
Calvert_Bluff <- readOGR(dsn = path, layer = "CB_400_original")
Simboro       <- readOGR(dsn = path, layer = "SB_400_original")
Hopper        <- readOGR(dsn = path, layer = "HP_400_original")
ShallowZone   <- readOGR(dsn = path, layer = "Total_400_original")

crs(ShallowZone) <- crs(Hopper)
## read in the spatial line df for river
path_river = ("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/hydrology")
#river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_GAM")
river <-readOGR(dsn = path_river, layer = "NHD_Flowline_Clp_perennial_clean_pts")
## read the DEM .tif file
DEM <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/rasters/DEM_POSGCD_NeighboringCounties.tif")

# import the pumping raster (difference in water level after pumping )
pumping500_2018 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2018.500ft.pst15.tif")
pumping500_2019 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2019.500ft.pst15.tif")
pumping500_2020 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")

pumping <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Pumping.Drawdown/Shallow.Combined.DrawDown.1929.2020.500ft.pst15.tif")
plot(pumping500_2018)
# import the GAM (simulated water levels) 
sim_WL2020 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2020.500ft.pst15.tif")
sim_WL2019 <- raster("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Data/500ft_res/Hds.Surfaces.Rasters.500ft.pst15/Shallow.Combined.Hds.2019.500ft.pst15.tif")
plot(sim_WL2020)
plot(sim_WL2019)


setwd("C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/R Code")

```

# create grids for kriging interpolations 
```{r}
# x <- seq(5800000, 6100000,400)
# y <- seq(19333000, 19633000,400)

x <- seq(5749040.9338320437818766, 6159040.9338320437818766,500)
y <- seq(19251409.7630450688302517, 19747409.7630450688302517,500)
data.grid <- meshgrid(x,y)

data.grid <- data.frame(unlist(matrix(data.grid$X)), unlist(matrix(data.grid$Y)))
colnames(data.grid) <- c("x","y")
data.grid$DEM = 0
coordinates(data.grid) <- ~ x + y
crs(data.grid) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
data.grid
```

# crop DEM topography raster to the same extent as the river and measured water levels 
```{r}
# create a empty raster to store topography DEM 
data.raster <- rasterFromXYZ(data.frame(data.grid@coords), res=c(500,500))
crs(data.raster) = CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
bb =raster::extent(5749040.9338320437818766, 6159040.9338320437818766, 19251409.7630450688302517, 19747409.7630450688302517) 
#c(5800000, 6100000,19333000,19633000)    #  6159040 - 5749040 = 410000 / 500 = 820   dimension of raster
data.raster[] = 0                                                                                      #  19747409 - 19251409 = 496000 / 500 = 992
data.raster

# DEM_raster <- raster(ncols = 750, nrows= 750)
# crs(DEM_raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
# extent(DEM_raster) <- extent(data.grid)
# DEM.raster <- resample(DEM, DEM_raster, method = "bilinear")
# DEM.raster
    
```

# Data wrangling, select specific cols for subsequent analysis and convert to spatial point data frame with correct CRS
```{r}
library(dplyr)
library(pracma)
# choose the year of WL you want to work on
data <- water_level %>%
  dplyr::select(Longitude, Latitude, SWN, DEM, Depth, ScrTop1, ScrBot1, ScrTop2, ScrBot2, SH_Note, Aq2Use, WL2019,WL2020) %>%
  # filter out NAs in both WL2019 and WL2020 to keep same number of measured water levels for both years
  filter(WL2019 != 'NA' & WL2020 != 'NA') 
head(data)

# we can also only exclude NAs for individual years
data_2019 <- data %>% 
  dplyr::select(Longitude, Latitude, WL2019) %>%
  filter(WL2019 != 'NA')

data_2020 <- data %>% 
  dplyr::select(Longitude, Latitude, WL2020) %>%
  filter(WL2020 != 'NA')

# create spatial point data frame with geographic coordinate systems
data_sp2019 <- SpatialPointsDataFrame(SpatialPoints(data_2019[,1:2]), data.frame(data_2019[,3]))
crs(data_sp2019) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

data_sp2020 <- SpatialPointsDataFrame(SpatialPoints(data_2020[,1:2]), data.frame(data_2020[,3]))
crs(data_sp2020) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

# create spatial point data frames with projected coordinate systems
new_crs <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs ")
data_sp2019 <- spTransform(data_sp2019,new_crs)
data_sp2020 <- spTransform(data_sp2020,new_crs)

# rename the colnames
colnames(data_sp2019@coords)[1:2] = c("x","y")
colnames(data_sp2019@data)        = c("WL2019")  
colnames(data_sp2020@coords)[1:2] = c("x","y")
colnames(data_sp2020@data)        = c("WL2020")  

# create combined spatial point data frame for both years (2019 & 2020) with projected coordinate systems
data_spadf <- SpatialPointsDataFrame(SpatialPoints(data[,1:2]),data=data[,12:13])
crs(data_spadf) = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
data_spadf <- spTransform(data_spadf,new_crs)
data_spadf@bbox <- t(matrix(extent(data.grid),2))  #crop
data_spadf


#Normalized Data
normalized = function (x) {(x-min(x))/(max(x)-min(x))}                              # create a normalization function
data_sp2019@data$WL2019n <- normalized(data_sp2019@data$WL2019)
data_sp2020@data$WL2020n <- normalized(data_sp2020@data$WL2020)
data_spadf@data$WL2019n  <- normalized(data_spadf@data$WL2019)
data_spadf@data$WL2020n  <- normalized(data_spadf@data$WL2020)

```


# Modelling the spatial structure of the target variable - make variogram
# 1. the variogram method computes the experimental variogram as a variogram object
# 2. the vgm method creates a variogram model object
# 3. the fit.variogram method adjusts a variogram model object to a variogram object
```{r}
# look at the spatial point data frame
data_spadf

# plot the empirical variogram for measured water levels at specific years
plot(v.WLn <- variogram(WL2019n ~ 1 , data=data_spadf, cutoff=300000, width=20000), main = "Empirical Variogram",pl = T)
# create a model for the empirical variogram
m.WLn <- vgm(0.05,"Gau",200000,0.003)
# plot both the fitted variogram model with the empirical variogram
plot(v.WLn, pl=T, model=m.WLn, main = "Variogram Model")
(m.WLn.f <- fit.variogram(v.WLn, m.WLn))

plot(v.WLn, pl=T, model=m.WLn.f, main = "Fitted OK Variogram Model")
```
## Ordinary Kriging Interpolation of the Target Variable
```{r}
# use krige function for oridinary kriging 
ko <- krige(WL2019n~1, locations=data_spadf, newdata=data.grid, model=m.WLn.f)
# extract the kriging estimates and kriging variances 
ko_pred <- data.frame(ko$var1.pred)
ko_var <- data.frame(ko$var1.var)

# use spplot to plot the kriging results 
sp::spplot(ko, "var1.pred", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Ordinary Kirging Estimates")
sp::spplot(ko, "var1.var", pch=19, col = "transparent",
     scales=list(draw=TRUE), main="Ordinary Kirging Estimates")

```

# plot contours
```{r}
ko.dataframe = as.data.frame(ko)
pred <- ggplot(ko.dataframe, aes(x ,y, z = var1.pred)) +
  geom_contour_filled(bins = 20)
var <- ggplot(ko.dataframe, aes(x ,y, z = var1.var)) +
  geom_contour_filled(bins = 20)
pred
```

# extract values for each acquifer
```{r}
# extract values of each spatial polygon from the kriging estimates, which is a spatial point data frame
Yegua_Jackson_df      <- as.data.frame(ko[Yegua_Jackson,])
Sparta_df             <- as.data.frame(ko[Sparta,])
Queen_City_df         <- as.data.frame(ko[Queen_City,])
Carrizo_df            <- as.data.frame(ko[Carrizo,])
Calvert_Bluff_df      <- as.data.frame(ko[Calvert_Bluff,])
Simboro_df            <- as.data.frame(ko[Simboro,])
Hopper_df             <- as.data.frame(ko[Hopper,])
ShallowZone_df        <- as.data.frame(ko[ShallowZone,]) 
# find the min and max of the original scale measured water levels
min <- min(data_spadf@data$WL2020)
max <- max(data_spadf@data$WL2020)
# back transform to original scales
InvNormalize <- function (x) {x * (max - min) + min}

# find one average value for each acquifer on the original scale using InvNormalize function
tol_avg   <- mean(InvNormalize(ko$var1.pred))
YJ_avg    <- mean(InvNormalize(Yegua_Jackson_df$var1.pred))
SPA_avg   <- mean(InvNormalize(Sparta_df$var1.pred))
QC_avg    <- mean(InvNormalize(Queen_City_df$var1.pred))
CAR_avg   <- mean(InvNormalize(Carrizo_df$var1.pred))
CB_avg    <- mean(InvNormalize(Calvert_Bluff_df$var1.pred))
SIM_avg   <- mean(InvNormalize(Simboro_df$var1.pred))
HOP_avg   <- mean(InvNormalize(Hopper_df$var1.pred))
SHAZ_avg  <- mean(InvNormalize(ShallowZone_df$var1.pred))
OK_avg <- data.frame(name = c("YJ", "SPA", "QC", "CAR", "CB", "SIM", "HOP", "TOL", "SHAZ"), avg = c(YJ_avg, SPA_avg, QC_avg, CAR_avg, CB_avg, SIM_avg, HOP_avg, tol_avg, SHAZ_avg))
OK_avg

```



# create raster and write on file
```{r}
# specify the image path on disk
image_path = "C:/Users/12819/Dropbox/My PC (DESKTOP-9VC04LJ)/Documents/Research/Ground Water/Output Shapefile"
# create raster and store it 
# only extract the predicted water levels 
ko@data$var1.predN <- InvNormalize(ko$var1.pred)
ko <- ko[,-(1:2)]

ko
# ko.raster <- raster(nrows = 10000,ncols =10000)
# crs(ko.raster) <- CRS("+proj=aea +lat_0=31.25 +lon_0=-100 +lat_1=27.5 +lat_2=35 +x_0=1500000 +y_0=6000000 +datum=NAD83 +units=us-ft +no_defs")
# extent(ko.raster) <- extent(ko)
# ko.raster <- rasterize(ko, ko.raster, field = "var1.predN", fun = mean)
# ko.raster

# write the kriging estimates as a spatial point data frame and save as a shape file 
writeOGR(obj = ko, layer = "OK_2019", dsn = image_path, driver = "ESRI Shapefile", overwrite_layer = TRUE)

```

















